---
title: "Final Project"
author: "Hao He"
date: "12/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(survey)
getwd()
setwd("C:/Users/hehao/Downloads/506_final/final_project/data")
```

# Introduction  

Sleeping time constitutes almost 1/3 of our daily activities. It is closely 
related to people's mood, working efficiency and health conditions. In this 
analysis, we will consider blood pressure as an indicator of people's health
condition and we aim to investigate the following question:   

> Do people with sleeping disorder tend to have higher blood pressures? 

**R** is the main software we will use for this analysis. Packages *survey*
and *dplyr* are also used.  

We will proceed this analysis by defining binary variables that distinguish
people with/without sleep disorder. Then we will provide point estimates and
95% confidence intervals for the mean blood pressure for people with/without
sleep disorder. We will also find points estimates and confidence intervals 
for the difference between these two groups of people and check whether 0 is
included in the confidence interval. Finally, we will use a regression 
approach to study whether sleep disorder is a significant factor that 
determine blood pressures or not.

# Data
We will use the 2017-2018 NHANES [demographic](https://wwwn.cdc.gov/Nchs/Nhanes/2017-2018/DEMO_J.XPT),
[sleep disorder](https://wwwn.cdc.gov/Nchs/Nhanes/2017-2018/SLQ_J.XPT) and
[blood pressure](https://wwwn.cdc.gov/Nchs/Nhanes/2017-2018/BPX_J.XPT) data
for this analysis.    
The specific variables of interests are listed below.  
`SEQN`: id  
`RIDAGEYR`: age
`RIAGENDR` : gender
`RIDSTATR`: status  
`SDMVPSU`: pseudo-PSU  
`SDMVSTRA`: pseudo-stratum  
`WTMEC2YR`: MEC exam weight  
`SLD012`:weekday sleep hours  
`SLD013`:weekend sleep hours  
`SLQ050`:ever told doctor had trouble sleeping  
`SLQ120`:how often fell overly sleepy during day  
`BPXSY1`:Systolic Blood Presure 1st measurement  
`BPXSY2`:Systolic Blood Presure 2nd measurement  
`BPXSY3`:Systolic Blood Presure 3rd measurement  
`BPXDI1`:Diastolic Blood Presure 1st measurement  
`BPXDI2`:Diastolic Blood Presure 2nd measurement  
`BPXDI3`:Diastolic Blood Presure 3rd measurement  

# Method

### Data Manipulation

```{r load data and some manipulation, include = FALSE}
setwd("C:/Users/hehao/Downloads/506_final/final_project/data")
# Load data and do some manipulation
demo = haven::read_xpt("DEMO_J.XPT") %>% 
  select(id = SEQN, status = RIDSTATR, weight = WTMEC2YR, age = RIDAGEYR,
         gender = RIAGENDR, psu = SDMVPSU, strata = SDMVSTRA) %>%
  filter(status == 2)
bp = haven::read_xpt("BPX_J.XPT") %>%
  select(id = SEQN, BPXSY1, BPXSY2, BPXSY3, BPXDI1, BPXDI2, BPXDI3) %>%
  filter(BPXSY1 >= 0 & BPXSY2 >=0 & BPXSY3 >= 0) %>%
  filter(BPXDI1 >= 0 & BPXDI2 >=0 & BPXDI3 >= 0) %>%
  mutate(BPXSY = (BPXSY1 + BPXSY2 + BPXSY3)/3) %>% 
  mutate(BPXDI = (BPXDI1 + BPXDI2 + BPXDI3)/3)
sleep = haven::read_xpt("SLQ_J.XPT") %>%
  select(id = SEQN, wkday_hour = SLD012, wkend_hour = SLD013,
         told_doctor = SLQ050, over_sleepy = SLQ120) %>%
  filter(wkday_hour < 15 & wkend_hour < 15 & told_doctor < 7 & 
           over_sleepy < 5) %>%
  mutate(is_disorder = ifelse(wkday_hour < 6 |
                                wkend_hour < 7 |
                                told_doctor == 1 |
                                over_sleepy > 3,1,0))
merged = demo %>%
  inner_join(sleep, by = "id") %>%
  inner_join(bp, by = "id") %>%
  select(id, age, gender, weight, psu, strata, is_disorder, BPXSY, BPXDI) %>%
  mutate(gender = ifelse(gender == 1, "Male", "Female"))
merged$gender = as.factor(merged$gender)
```

We will create the following variables to represent the subjects' systolic 
and diastolic blood pressure level, and whether the subject has sleep disorder
or not.

`BPXSY` : the average systolic blood pressure out of three measurements  

`BPXDI` : the average diastolic blood pressure out of three measurements  

`is_disorder` *= 1*  if any of the following conditions is met: weekday sleep
hour less than 6, weekend sleep hour less than 7, have ever told doctor with
trouble sleeping, feel overly sleepy for more than 15 times per month. And *= 0* otherwise.

Codes for the data manipulation stated above and all the analysis below 
can be found at 
[github_link](https://github.com/hhaohe513/final_project) in final_project.R 

### Point Estimate and Confidence Interval
We will use the survey
design([svydesign()](https://cran.r-project.org/web/packages/survey/survey.pdf))
to find accurate point estimates and confidence intervals for each
group(with/without sleep disorder) and also their differences.

The codes for using svydesign() is presented below:  

```{r svydesign code, include = TRUE}
svy = svydesign(ids = ~psu, weights = ~weight, strata = ~strata,
                data = merged, nest = TRUE)
mean_se = as.data.frame(svyby(~BPXSY + BPXDI,
                              by = ~is_disorder, svy,  svymean))
```

### Regression Approach

For the regression approach, we will use the svyglm() function to fit linear
models for both systolic and diastolic blood pressure. We will consider 
systolic/diastolic blood pressure as response variables and age, gender and
the indicator variable is_disorder as predictors. 

Since the main goal of our analysis is to make inference about the relationship
between blood pressure and sleep disorder, we will focus on the parameter
coefficients and significance instead of model performances. 

The formulas below indicate the models we are fitting.  

```{}
BPXSY ~ gender + age + is_disorder
BPXDI ~ gender + age + is_disorder
```



# Results

### Result from point estimate and confidence interval
```{r point estimate results, include = FALSE}
nice_table = data.frame("Variables" = c("BPXSY","BPXDI"),
                        "mean_disorder" = c(mean_se[2,2],mean_se[2,3]),
                        "mean_normal" = c(mean_se[1,2],mean_se[1,3]),
                        "se_disorder" = c(mean_se[2,4],mean_se[2,5]),
                        "se_normal" = c(mean_se[2,4],mean_se[2,5])) %>%
  mutate(mean_diff = mean_disorder - mean_normal) %>%
  mutate(lwr_diff = mean_diff - sqrt(se_disorder^2 + se_normal^2)*qnorm(0.975),
         upr_diff = mean_diff + 
           sqrt(se_disorder^2 + se_normal^2)*qnorm(0.975)) %>%
  select(Variables, mean_disorder, mean_normal, mean_diff, lwr_diff, upr_diff)
```

```{r table for presentation, include = TRUE}
nice_table %>% 
  knitr::kable(
    format = 'html', 
    digits = 4
  ) %>%
  kableExtra::kable_styling("striped", full_width = TRUE)
```

From the resulting table, we see that for both systolic and diastolic blood
pressure, 0 is not included in the confidence interval for the difference
between groups with sleep disorder and groups without sleep disorder. Note that
the lower bound for the difference in diastolic blood pressure is very close to
0. This approach only considers mean estimation, which is not suffice for 
further inference. So we will consider regression approach using *svyglm*
function.

### Resluts from regression approach
```{r regression fit,include = FALSE}
sys_glm = svyglm(BPXSY ~ gender + age + is_disorder, design = svy)
dias_glm = svyglm(BPXDI ~ gender + age + is_disorder, design = svy)
```

```{r regression result, include = TRUE}
summary(sys_glm)
summary(dias_glm)
```

From the model result, we observe that for both models, sleep disorder is a
significant predictor in the model. The coefficients from the model results 
suggest that people with sleep disorder have 2.05 higher systolic blood
pressure and 1.46 higher diastolic blood pressure than those who don't have
sleep disorder. These numbers are slightly different from the estimated
difference because we include age and gender as covariates.

# Conclusion  

The results from point estimation and regression coincides with each other, so
we are confident to conclude that people with sleep disorder tend to have 
higher blood pressure than those who don't. However, we must notice that this
analysis is very simplistic and limited. The regression approach could be done
in much more complicated ways by including more variables and considering more
possible models. With more options of models, we may use methods like 
cross-validation to test for model performances and do model selection. The 
interpretation may also become more interesting by discussing the potential 
confounding variables and the collider effects among variables.

